### 2021-03-25 manuscript_00.Rmd
text updates

---
title: 'Building the Meta Single Cell Transcriptome Ocular Meta-Atlas'
author:
  - Vinay Swamy:
      institute: bg
  - Temesgen Fufa:
      institute: mgog
  - Robert Hufnagel:
      institute: mgog
  - David McGaughey:
      institute:
        - bg
      correspondence: "yes"
      email: mcgaugheyd@mail.nih.gov
institute:
  - bg: Bioinformatics Group, Ophthalmic Genetics & Visual Function Branch, National Eye Institute, National Institutes of Health
  - mgog: Medical Genetics and Ophthalmic Genomics Unit, National Eye Institute, National Institutes of Health

date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  word_document:
    reference_docx: word-styles-reference-01.docx
    fig_caption: yes
    pandoc_args:
      - '--lua-filter=scholarly-metadata.lua'
      - '--lua-filter=author-info-blocks.lua'
bibliography: references.bib
csl: investigative-ophthalmology-and-visual-science.csl
abstract: "The development of highly scalable single cell transcriptome technology has resulted in the creation of thousands of datasets, over 30 in the retina alone. Analyzing the transcriptomes between different projects is highly desirable as this would allow for better assessment of which biological effects are consistent across independent studies. However it is difficult to compare and contrast data across different projects as there are substantial batch effects from computational processing, single cell technology utilized, and the natural biological variation. While many single cell transcriptome specific batch correction methods purport to remove the technical noise it is difficult to ascertain which method functions works best. We developed a lightweight R package (scPOP) that brings in batch integration methods and uses a simple heuristic to balance batch merging and celltype/cluster purity. We use this package along with a Snakefile based workflow system to demonstrate how to optimally merge 766,615 cells from 34 retina datsets and three species to create a massive ocular single cell transcriptome meta-atlas. This provides a model how to efficiently create meta-atlases for tissues and cells of interest."
keywords: "RNA-seq, retina, RPE, ocular, eye, Snakemake, singlecell, scRNA, single-cell, singlecell, scPOP, R"
---

```{r Setup, message=FALSE, warning=FALSE, include=FALSE}
#knitr::opts_chunk$set(fig.pos = 'p') # Places figures on their own pages
knitr::opts_chunk$set(out.width = '100%', dpi=300)
library(tidyverse)
library(citr)
library(cowplot)
library(ggrepel)
library(colorspace)
library(flextable)
library(captioner)
library(pool)
library(RSQLite)
library(formattable)
library(ggalluvial)
library(scattermore)
library(grid)
library(ComplexHeatmap)
library(viridis)
library(Matrix)
library(MatrixGenerics)
library(matrixStats)
library(velociraptor)
library(SingleCellExperiment)
# setup caption-ing
fig_cap <- captioner("Figure")
supFig_cap <- captioner("Supplemental Figure")
#tab_cap <- captioner("Table")
supTab_cap <- captioner("Supplemental Table")

scEiaD_2020_v01 <- dbPool(drv = SQLite(), dbname = "~/data/scEiaD/MOARTABLES__anthology_limmaFALSE___5000-transform-counts-universe-batch-scVIprojectionSO-8-0.1-50-5.sqlite", idleTimeout = 3600000)
x_dir <- -1
y_dir <- 1

meta_filter <- fst::read_fst('~/data/scEiaD/2021_03_17_meta_filter.fst') %>% as_tibble()
meta_filter <- meta_filter %>% mutate(CellType_predict = case_when(CellType_predict == 'Photoreceptor Precursors' ~ 'PR Precursors',
                                                                   CellType_predict == 'AC/HC_Precurs' ~ 'AC/HC Precursors',
                                                                   CellType_predict == 'RPC' ~ 'RPCs',
                                                                   CellType_predict == 'Mesenchymal/RPE/Endothelial' ~ 'Endothelial',
                                                                   TRUE ~ CellType_predict),
                                      CellType = case_when(CellType == 'Photoreceptor Precursors' ~ 'PR Precursors',
                                                           CellType == 'AC/HC_Precurs' ~ 'AC/HC Precursors',
                                                           CellType == 'RPC' ~ 'RPCs',
                                                           CellType == 'Mesenchymal/RPE/Endothelial' ~ 'Endothelial',
                                                           TRUE ~ CellType)) %>%
  mutate(UMAP_a = UMAP_2 * x_dir,
         UMAP_b = UMAP_1 * y_dir) %>%
  mutate(UMAP_1 = UMAP_a, UMAP_2 = UMAP_b) %>%
  mutate(CellType_predict = case_when(!is.na(TabulaMurisCellType_predict) ~ 'Tabula Muris',
                                      TRUE ~ CellType_predict))

# meta data before doublet removal
load('~/data/scEiaD/n_features-5000__transform-counts__partition-universe__covariate-batch__method-scVIprojectionSO__dims-8__preFilter.scEiaD__dist-0.1__neighbors-50.umap.Rdata')
umapRef <- umap

methods <- factor(c('scArches', 'bbknn','insct','magic', 'scVI','CCA', 'scanorama', 'harmony', 'fastMNN', 'combat', 'none'), levels = c('bbknn','CCA','combat','fastMNN','harmony','insct','magic','scanorama','scArches','scVI', 'none'))

gse <- read_tsv('~/git/scEiaD/data/GEO_Study_Level_Metadata.tsv') %>% filter(study_accession %in% meta_filter$study_accession)

srt <- data.table::fread('~/git/scEiaD/data/sample_run_layout_organism_tech.tsv')
qc <- data.table::fread('~/data/scEiaD/QC.tsv.gz')
qc <- qc %>% left_join(srt %>% select(sample_accession, SX = Source) %>% unique(), by = 'sample_accession') %>% filter(SX %in% c('iPSC','Tissue'))
# cell type prediction
load('~/data/scEiaD/predictions_n_features-5000__transform-counts__partition-universe__covariate-batch__method-scVIprojectionSO__dims-8__preFilter.scEiaD__dist-0.1__neighbors-50.Rdata')

test_predictions <- model_out$test_dat %>% as_tibble %>% 
  mutate(max_pred_prob = rowMaxs(.[,-(31:34)] %>% as.matrix), 
         pred_correct = ifelse(CellTypeID == true_label_id, 'Correct', 'incorrect'))
# created by running 'source' on src/make_dotplot_markers.R
load('~/data/scEiaD/top_markers.Rdata')
load('~/data/scEiaD/top_marker_pmid.Rdata')

# replace cone bipolar with bipolar
meta_filter$CellType <- gsub('Cone Bipolar','Bipolar', meta_filter$CellType)

# kallisto stats
run_info <- read_tsv('~/data/scEiaD/aggregated_run_info.tsv.gz')
write_tsv(run_info %>% 
            filter(sample_accession %in% meta_filter$sample_accession) %>% 
            select(-file) %>% 
            left_join(meta_filter %>% 
                        group_by(sample_accession) %>% 
                        summarise(nCount_RNA = mean(nCount_RNA), nFeature_RNA = mean(nFeature_RNA), percent_mt = mean(percent_mt)) %>% 
                        select(sample_accession, nCount_RNA, nFeature_RNA, percent_mt)), 
          file = 'supplemental_file__kallisto_stats.tsv')

# write top markers to sup file
write_tsv(top_markers, file = 'supplemental_file__celltype_markers.tsv')

# load mini human velocity
load('~/data/scEiaD/velo_scEiaD_2021-03-18_basemodel_homosapiens_tiny.Rdata')
umap_mat <- reducedDim(velo_sceX, 'X_scviUMAP')
grid.df <- gridVectors(umap_mat, reducedDim(velo_sceX, 'velocity_scviUMAP'))

# load ML species transfer data
# created with src/organism_celltype_ML.R
load('~/data/scEiaD/species_transfer_celltype_model.Rdata')
```

# Introduction

## A plethora of single-cell transcriptome studies in the retina

The retina contains a multitude of cell types that, in total, are responsible for turning light information into signal for the brain to interpret as vision. Very briefly, the photoreceptors (rods and cones) are responsible for capturing the photons. The retinal pigmented epithelium (RPE) behind the photoreceptors physically support the rods and cones by processing byproducts of the visual cycle. Müller glia serve as support cells for the neurons. The retinal bipolar cells transmit the electrical signal from the photoreceptors to the retinal ganglion cells. Horizontal and amacrine regulate and help interpret signals from the photoreceptors. The signal is relayed via the retinal ganglion projections through the optic stalk to the brain (for review see [@maslandNeuronalOrganizationRetina2012]). Since 2000 many groups have investigated gene expression in small numbers of individual cells of the retina [@anniesIsoformPatternAChR2002a; @hagstromConePigmentGene2000; @trimarchiMolecularHeterogeneityDeveloping2007a; @wahlinMethodAnalysisGene2004].

The  recent introduction of lower cost and high throughput single cell sequencing technology has led to an explosion of research across many fields. As of early 2021, over 40 million cells have been sequenced across over 1,200 studies and the average size of each study starting in 2020 is over 100,000 cells [@svenssonCuratedDatabaseReveals2020]. The retina was used as the source tissue in one of the earliest works in the high throughput single cell transcriptomics field [@macoskoHighlyParallelGenomewide2015]. As of late 2020, over a twenty published studies, cumulatively containing over a million cells, have used single cell technology to profile cell type specific gene expression patterns, cell fate trajectory, tissue and cell differentiation, and disease perturbation across multiple mammalian species [@buenaventuraIdentificationGenesEnriched2019; @clarkComprehensiveAnalysisRetinal2018; @collinCRXExpressionPluripotent2019; @collinDeconstructingRetinalOrganoids2019; @cowanCellTypesHuman2020; @diacouSix3Six6Are2018; @eldredThyroidHormoneSignaling2018; @hengHypoxiaToleranceNorrindeficient2019; @huDissectingTranscriptomeLandscape2019; @logiudiceSinglecellTranscriptionalLogic2019; @lukowskiSinglecellTranscriptomeAtlas2019; @luSingleCellAnalysisHuman2020;  @menonSinglecellTranscriptomicAtlas2019; @okorenMicroglialFunctionDistinct2019; @pengMolecularClassificationComparative2019; @phillipsNovelApproachSingle2018; @rheaumeSingleCellTranscriptome2018; @ruedaHippoPathwayBlocks2019; @shekharComprehensiveClassificationRetinal2016; @sridharSingleCellTranscriptomicComparison2020; @tranSingleCellProfilesRetinal2019; @voigtMolecularCharacterizationFoveal2019; @voigtSinglecellTranscriptomicsHuman2019; @yanCellAtlasHuman2020].

<!-- ## Many researchers use processed count tables for methods development and hypothesis testing -->

While the gene - cell count tables are generally made available in repositories like the Gene Expression Omnibus (GEO), there are no requirements to uniformly process the data. This means the count tables cannot be used in cross-study comparisons as even small differences in the computational pipeline (aligner, transcriptome reference, etc.) create study-specific effects. This issue can be addressed only by re-quantifying the data in a uniform computational environment. Fortunately, due to the continued development of computationally light-weight gene quantification tools in the single-cell space (e.g kallisto bustools, alevin-fry), re-quantification does not require massive compute and time resources [@melstedModularEfficientPreprocessing2019; @srivastavaAlevinEfficientlyEstimates2019].

Still, even after re-quantification under identical computational conditions there remain study specific batch effects due to the diversity in single cell technologies used and variation in tissue handling and processing across each scientific group. The single cell community has recognized that removal of these technical (also referred to as batch) effects is a critical issue and have independently developed many tools, though it remains unclear which tools and parameters are optimal for a particular dataset [@butlerIntegratingSinglecellTranscriptomic2018; @haghverdiBatchEffectsSinglecell2018; @hieEfficientIntegrationHeterogeneous2019; @johnsonAdjustingBatchEffects2007; @korsunskyFastSensitiveAccurate2019; @liuJointlyDefiningCell2020; @lopezDeepGenerativeModeling2018; @polanskiBBKNNFastBatch2020; @QueryReferenceSinglecell; @simonINSCTIntegratingMillions2020; @stuartComprehensiveIntegrationSingleCell2019; @vandijkRecoveringGeneInteractions2018].

<!-- Study specific web portals, should they exist, are silos which only contain the single cell transcriptomes related to their study. There are notable efforts to create web portals (UCSC Single Cell Browser, EBI Single Cell Expression Atlas) that allow for GUI-based viewing of data for most publicly available datasets. However, these again have the shortcoming as the study-specific web portals discussed above in that you still cannot directly compare between independent studies. In these situations it is impossible to compare directly across datasets. When exploring a gene expression pattern, it is crucial to have reproducibility across disparate groups to enhance confidence that effect is biological as bench experiments require large investments in time.  -->

<!-- ## Aggregation of results to create meta-atlas important -->

<!-- While these groups has leveraged this new technology to [do stuff], there is a large chasm to cross the enable outside groups to use the data in new ways. Several of these papers are accompanied by reactive web apps, which allow for relatively quick checks of gene expression across cell type or cluster assignment. However these web apps have minimal to no data exploration tools and are slow to operate, especially with the larger datasets.  -->

<!-- While lightly processed counts data is often provided, differences in both single cell technology (e.g. droplet or well based) and bioinformatic processing choices make it impossible to simply concatenate the datasets together as differences found will likely be technical.  -->

## The projectable meta-atlas

We propose that by re-processing publicly available raw single cell transcriptome data in a consistent bioinformatic framework and optimally using batch correction tools we can create a meta-atlas of retina single cell transcriptomes. As there are thousands of possible permutations of single cell tools, references, and parameter choices, we create our meta-atlas (which we refer to as the single cell Eye in a Disk or scEiaD) by benchmarking integration outcomes across multiple important single cell RNA-seq processing parameters (batch removal method, number of hyper-variable genes (HVGs), clustering resolution, etc.). The benchmarking system we developed uses a wide variety of metrics that combine in the R package scPOP (single-cell Pick Optimal Parameters). The scEiaD will be of utility to two communities. First, the ocular community who can both search scEiaD for gene expression across many dimensions (e.g. cluster, cell type, study) and project their own single cell data onto scEiaD for comparison and rich automatic cell labelling. Second, the computational community can use this very large, well-curated dataset to test algorithms for compute efficiency and performance in a diverse environment. As we believe data re-use is a powerful and efficient approach to facilitate discovery, we provide our meta-atlas code-base, the meta-atlas in several data formats, and propose general guidelines to optimally create custom meta-atlases.

# Results

## We identify `r meta_filter %>% pull(study_accession) %>% unique() %>% length() - 1` ocular scRNA datasets across 3 species

The first step in building a meta-atlas is identifying studies to draw the data from. We identified ocular single cell RNA sequencing (scRNA) studies by querying PubMed, the Sequence Read Archive (SRA), and the European Nucleotide Archive (ENA) for the inclusive terms "retina", "single cell", "scRNA", "ocular", "eye", "transcriptome." We then hand filtered the results to only keep ocular and normal (non-perturbed or mutagenized) data from single cell RNA-seq technology. On December 2020 we identified `r meta_filter %>% pull(study_accession) %>% unique() %>% length() - 1` deposited datasets that have been published in `r meta_filter %>% pull(Citation) %>% unique() %>% length()` publications (`r fig_cap(name = 'fig1', display = 'cite')`). To provide a non-ocular reference we also downloaded the raw sequence data from the Tabula Muris project for re-processing. In cases where the fastq file from the SRA was not processed properly, we acquired the bam files (SRA or personal correspondence) and re-extracted the fastq. After downloading all the data we had 11 TB across `r scEiaD_2020_v01 %>% tbl('metadata_filter') %>% pull(sample_accession) %>% unique() %>% length()` fastq file sets.

```{r fig1, fig.width=12, fig.height=6, fig.cap=fig1_cap, echo=FALSE, message = FALSE, warning = FALSE}
fig1_cap <- fig_cap(name='fig1', caption =  'a. Schematic of the retina with major cell types delineated b. Simplified directed workflow of major steps in scEiaD creation from raw counts to gene counts, benchmarking optimal integration methods (SnakePOP) to produce batch corrected latent dimensions (Latent Dims), then downstream analysis outputs like clustering, differential gene testing (Diff Testing), and 2D UMAP visualization. c. Counts of published papers and batches (unique biological samples) for each scRNA technlogy, split by organism d. Cell type counts extracted from published studies for the more common retina cell types, split by species. Count of study accessions for each species overlaid on bar plot.')

source('figs_and_tables/fig1.R')
retina_cartoon <- ggdraw() + draw_image("figs_and_tables/retina_cells_flat2.png", scale = 1)
plot_grid(
  plot_grid(retina_cartoon, b + coord_cartesian(clip = "off"), nrow = 1, labels = c('','b'), rel_widths = c(0.5,1)), 
  plot_grid(a, c, align = 'v',  axis = 'lr', nrow = 1, labels = c('','d')),
  nrow = 2, 
  rel_heights= c(1,1), 
  labels = c('a','c'))
```

## Transcriptome quantification across multiple technologies
Droplet and well based scRNA-seq technologies require different quantification approaches as the former have UMI and multiple cells are quantified within a single file. We wrote a Snakemake based pipeline (SnakeQUANT, see methods) to quantify and merge both droplet and well based technologies into a single matrix for downstream processing.  For well based data we perform both gene and isoform level quantification; for droplet base technologies we quantify both exonic and intronic gene-level expression to facilitate calculation of RNA velocity. In total we quantify `r run_info %>% filter(sample_accession %in% meta_filter$sample_accession) %>% pull(n_processed) %>% sum() %>% formatC(., digits = 2)` molecules, finding `r run_info %>% filter(sample_accession %in% meta_filter$sample_accession) %>% pull(n_unique) %>% sum() %>% formatC(., digits = 2)` unique molecules with a mean pseudoalignment rate of `r ((run_info$n_pseudoaligned %>% sum() ) / (run_info$n_processed %>% sum()) * 100)  %>% formatC(., digits = 3)`%. Across the `r meta_filter %>% nrow() %>% format(., big.mark=",", scientific=FALSE)` cells (post QC) we have an average of `r meta_filter$nCount_RNA %>% mean() %>% formatC(., digits = 4, big.mark=",")` RNA counts across `r meta_filter$nFeature_RNA %>% mean() %>% formatC(., digits = 3, big.mark=",")` unique genes (see Supplemental File kallisto_stats.tsv and splicing_stats.tsv for more details).

## `r qc %>% nrow() %>% as.integer() %>% format(., big.mark=",", scientific=FALSE)` cells before quality control

Gene-level counts were quantified with the kallisto bustools pseudo-aligner for both the droplet and well based samples. After empty droplet removal, we had `r format(qc %>% nrow() %>% as.integer() , big.mark=",", scientific=FALSE)` cells. We then removed cells which had more than 10% mitochondrial reads across all gene counts or fewer than 200 unique genes quantified. After these standard quality control steps we were left with `r format(umapRef %>% nrow, big.mark = ",", scientific=FALSE)` cells (`r supFig_cap(name = 'supFig1', display = 'cite')`).

A core objective of many scRNA based studies is labeling the cell types. As this information is crucial to assess dataset integration and provide an accurate reference for user querying, we extracted individual cell labels with a combination of inspecting the GEO web site, supplemental information from the publication, web resources (e.g. a web app was created for the paper), and personal correspondence. After normalizing cell type name nomenclature, we obtained labels for `r meta_filter$CellType %>% table() %>% sum() %>% format(., big.mark=",", scientific=FALSE)` cells across `r meta_filter$CellType %>% table() %>% length() ` cell types (`r supTab_cap(name = 'supTab1_CTcounts', display = 'cite')`).

## Running `r methods %>% length()` tools in a Snakemake-based system

Disentangling the technical and biological effects when integrating multiple datasets is crucial. We define batch as each unique biological sample and assume each study is at least one unique sample. We studied the metadata and methods of each study to identify the unique biological samples. Within the current scEiaD data set we identified `r umapRef %>% pull(batch) %>% unique() %>% length()` batches across `r umapRef %>% pull(study_accession) %>% unique() %>% length()` deposited datasets in `r umapRef %>% left_join(gse) %>% pull(PMID) %>% unique() %>% length()` published papers.

A wide variety of methods have been written for scRNA-seq integration. As we were uncertain which would perform the best, we ran `r methods %>% length()` tools with a commonly used set of key parameters like number of hyper variable genes (HVG), size of latent dimensional space, and the number of nearest neighbors for the louvain clustering algorithm. The Snakemake system was used to automate the running of the wide variety of tools. In total 5,591 jobs were run to quantify gene expression and build the unified Seurat objects (SnakeQUANT) and 2,446 jobs were run to assess integration performance (SnakePOP).

The two key metrics which have to be balanced in order to optimize integration performance are cell type or cluster purity (where different cell types or clusters should be homogenous) and batch mixing (the same cell types should be similar across independent studies). While these can be visually assessed by looking at marker gene expression across the 2D UMAP projection, it is more rigorous and scalable to quantify these diametrically opposed characteristics.

## scPOP wraps several different methods for measuring integration performance.

Multiple methods have been proposed to quantitatively evaluate batch correction. Some of these metrics evaluate the concordance between sets of labels, while other compute distances between the individual data points of a given set of labels. While any one of these methods can be useful, we propose that calculating and evaluating them in tandem provides greater accuracy for dataset integration. We developed the R package scPOP, a lightweight, low dependency R package which brings together the Local Simpson Index (LISI), Average Silhouette Width (ASW), Adjusted Rand Index (ARI) and Normalized Mutual Information (NMI) metrics from the R packages Harmony, kBET and aricode, respectively. The LISI and ASW were used to measure batch mixing (where lower is better), cell type mixing (higher is better), and cluster mixing (higher is better). NMI and ARI were used to assess the consistency of cell type to cluster assignment (where 1 is perfect correspondence between cluster and cell type).

To visualize the interplay between batch mixing and cell type distinction we plot the batch mixing LISI score (which has been multiplied by -1) on the y-axis (higher is better) against the cluster LISI on the x-axis (higher is better). The best performer on both metrics will be in the top right corner  (`r supFig_cap(name = 'supFig_benchmark', display = 'cite')`a). In the same manner we plot the silhouette metric (`r supFig_cap(name = 'supFig_benchmark', display = 'cite')` b). To merge the different scores we define $sumZScale = \sum scale(m)$ where m is a metric (LISI by batch, LISI by cluster, LISI by cell type, silhouette by batch, silhouette by cluster, silhouette by celltype, NMI, and ARI).

On one extreme we have ComBat, which merges together different batches very well, but also mixes together the distinct cell types (`r fig_cap(name = 'figBenchmark', display = 'cite')`a). The other extreme is not using any batch integration method, where you see very distinct groups of cells, but also each nearly study is has a distinct region in the UMAP (`r fig_cap(name = 'figBenchmark', display = 'cite')`b). In our scEiaD dataset we see that like ComBat, Harmony and CCA are weighed more towards batch mixing then cluster and cell type purity. Scanorama, fastMNN, and bbknn prioritize cleanly separating the clusters. With our scEiaD meta-atlas insct, trVAE, and magic do not perform particularly well in batch mixing or cluster purity. Across the 11 integration methods we tested, scVI achieved the highest sumZScale for out scEiaD meta-atlas.

## Different normalization methods alter integration performance

<!-- https://www.frontiersin.org/articles/10.3389/fgene.2020.00041/full -->
There are several normalization approaches that have been used or published. The "standard" approach that the popular analysis packages Seurat and anndata use by default is to, per cell, divide the counts by the sum counts for the cell, multiply by a scaling factor, then log transform. This helps make the count distribution more normal, which is an assumption that many algorithms require. In contrast, the scran normalization method groups cells into pools and normalizes across the pool summed counts instead of the individual cell counts. We also use the square root (sqrt) normalization which replaces the log transformation with a square root. Library size (libSize) normalization omits the sqrt or log transformation. Finally some methods, like scVI, directly use the raw counts data for modeling the data.

As expected the libSize normalization which omits the log or square root scaling generally performs the worst (`r fig_cap(name = 'figBenchmark', display = 'cite')`c). We see that the remaining normalization techniques alter the batch correction performance, though the outcome differs across the different methods. We also see that changing the number of latent dimensions (8 or 30) can occasionally dramatically change performance. These results demonstrate the importance of assessing performance in a rigorous manner across many parameters.

<!-- For users who have custom needs, the weights of each metrics can be customized in scPOP to prioritize batch mixing or cluster/cell type separation. -->

```{r figBenchmark, fig.width=14, fig.height=9, fig.cap=figBenchmark_cap, echo=FALSE, message = FALSE, warning = FALSE}
figBenchmark_cap <- fig_cap(name='figBenchmark', caption =  'a. Example of a method (combat) which has a high level of batch blending, but poor separation of cell types (colored by cell type). b. no batch correction cleanly separates cell types but does not mix batches (colored by study). c. sumZScale (higher is better) for each method across a variety of data normalizations. All methods shown here use 2000 HVG, louvain clustering, and 8 latent dimensions. Each color is a different method. d. Boxplot of 4 different clustering resolutions for across 1000 to 10000 HVG numbers and 4 to 50 scVI latent dimensions. Open boxes are using scVI-standard and gray boxes are scVI-projection (human reference with the remaining data projected) ')

source('figs_and_tables/fig_umap_benchmarking.R')
plot_grid(
  plot_grid(
    umap_plot_maker(umapCombat), 
    umap_plot_maker(umapNone, color_against = 'study_accession'),
    rel_widths = c(1.3, 1),
    labels = c('a','b')), 
  plot_grid(
    zscore_sum_all_methods, 
    zscore_droplet_scVI_optimize,
    labels = c('c','d')),
  nrow = 2,
  rel_heights = c(1, 0.9)
)
```

<!-- ## scVI has highest integration performance -->

<!-- scPOP finds that for our dataset scVI has the strongest performance, with a sum score of SOMETHING. The next highest performers are fastMNN, harmony, and scanorama. We do not claim that these results are general - rather we emphasize that for each (meta) atlas creation that scPOP or some similar unbiased approach where a range of several key parameters (e.g. HVG number) are tested in a quantitative framework.  -->

## Further optimization of scVI with grid search and projection

To find the best set of parameters for scVI in out dataset we did a grid search across key parameters: HVG, latent dimensions, and k nearest neighbors. Furthermore we used a recent advance in scVI capability (>= version 0.8.0) adapted from scArches that allows one to build a reference model and query or project the new cells onto it [@QueryReferenceSinglecell]. We refer to this projectable model as “scVI-projection", and the previous scVI model as “scVI-standard”.  We built a scVI-projected model trained on human cells and then projected the mouse and macaque data onto it. We then compared scVI-projection against the against the previous scVI-standard across all the previously mentioned parameters. Using scPOP we first saw that the scVIprojection approach generally performed better than running scVI with all of the data (`r fig_cap(name = 'figBenchmark', display = 'cite')`d). We found the optimal overall parameters to be 5000 HVG, 8 latent dimensions, and with 5 k-nearest neighbors for the cluster finding. We also varied the UMAP projection values of nearest neighbors and minimum distance to qualitatively pick a 2D projection, selecting a minimum distance of 0.1 and 50 nearest neighbors.

## High accuracy xgboost ML model built to label unknown cell types across all technologies

To further study cell type specific expression patterning we needed to label the `r format(umapRef %>% nrow() - meta_filter %>% filter(!is.na(CellType)) %>% nrow(), big.mark=",", scientific=FALSE)` unlabeled cells. Traditionally this is done by clustering the cells, then using cell type specific markers to label the clusters. However, as we had hundreds of thousands of expert labeled cells across `r umapRef %>% filter(!is.na(CellType)) %>% left_join(gse) %>% pull(PMID) %>% unique() %>% length()` publications (`r fig_cap(name = 'fig1', display = 'cite')`a) we built a xgboost-based machine learning model that used 2/3 of the labeled cells as a training set (see methods for more details) to train a cell type predictor for scEiaD input. The trained model was used to predict the cell type assignments for all cells in scEiaD. In this manner we both label most cells (cells which cannot assign a cell type with a probability above 0.5 were left unlabeled) and correct a small number of probable mislabels in the truth set (`r fig_cap(name = 'figXGboost_umap', display = 'cite')`a).

As a brief case study of our ML performance, we look at the cell type labels assigned to a Shekhar et al. study where they used SMART-seq on a retinal bipolar cell enriched population of cells [@shekharComprehensiveClassificationRetinal2016] . Even though our ML algorithm was trained only on droplet-based data, our algorithm labels most of this dataset as retinal bipolar cells, with the next most common cell types being amacrine. The same result was found by Shekhkar et al. (`r supFig_cap(name = 'supFigShekharSMART', display = 'cite')`). To more generally evaluate performance we use the precision recall (PR) curve, which visualizes the ability of the model to precisely label known cells at a given confidence. The area under the PR curve (AUC) summarizes the effectiveness of the model across different cell types, with 1 being the highest performance. The xgboost model can predict rods, bipolar cells, and Müller glia with near perfect performance (`r supFig_cap(name = 'supFig_PR', display = 'cite')`). Most of the remaining cell types can be predicted with an AUC over 0.9 `r supTab_cap(name = 'supTab_AUCxgboost', display = 'cite')`. Several of the precursor cell types (photoreceptor, AC/HC, and neurogenic) were labelled with lower confidence `r supTab_cap(name = 'supTab_CTmislabels', display = 'cite')`. The next most common labels for these cell types were either other precursor cells (e.g. 100 AC/HC Precursors were labeled as Neurogenic) or the adjacent terminal cell type (e.g. 254 photoreceptor precursors were labeled as cones). Other cell types that were challenging for the model to predict were the artery, choriocapillaris, and vein. Artery, choriocapillaris, and vein are constructed from endothelial cells and we find that for all three of these, endothelial was the second most common label. Overall, our xgboost based ML model shows strong accuracy across the major cell types of the retina (overall AUC of 0.98).

```{r figXGboost_umap, fig.width=18, fig.height=13, fig.cap=figXGboost_umap_cap, echo=FALSE, message = FALSE, warning = FALSE}
figXGboost_umap_cap <- fig_cap(name='figXGboost_umap', caption =  'a. Top genes that are differentially expressed across the major cell types of the retina (PR is short for Photoreceptor). Genes are colored by which cell type they are differentially expressed in. The dot size is proportional to the percentage of those cells that have detectable levels of the gene. The color of the dot is the log2 scaled CPM expression. b. 2D UMAP projection of scEiaD, colored by cell type (Tabula Muris data is gray). Arrows are scvelo RNA velocity. Longer arrows are cells with higher velocity (relatively more unspliced transcripts). c. Facet plot that demonstrates how each major cell type of the retina is contained within a distinct space. d. Confusion matrix of cell type prediction performance of our xgboost labeller between predicted (x axis) and known (y axis) using data withheld from the machine learner. Most of the cell types are indeed labelled as their true type. e. Faceting of 2D UMAP by species and colored by cell type demonstrate how the major cell types of the retina share space with like cell types, despite being from mouse, human, and macaque.')

source('figs_and_tables/sup_fig2__celltype_xgboost_labelling.R')
source('figs_and_tables/make_meta_scatter_umap_plot.R')
source('figs_and_tables/dotplot.R')

# markers for dotplot
input <- list()
input[['dotplot_Gene']] <- top_markers %>%  
  filter(grepl('Amacrine|Rod|Cone|Retinal|Muller|Horizon|Bipol|Astro|RPE', cluster),
         !grepl('C1orf|MT-', Gene)) %>%  
  group_by(cluster) %>% slice_max(mean_auc, n =5) %>% 
  pull(Gene)
# bipolar have none with auc_count > 0, so run again
# input[['dotplot_Gene']] <- c(input[['dotplot_Gene']], 
#                              top_markers %>%  
#   filter(cluster %in% c('Amacrine Cells', 'Bipolar Cells'), !grepl('C1orf|MT-', Gene), `log.p.adj` < -10000) %>%  
#   group_by(cluster) %>% slice_min(`log.p.adj`, n =5) %>% 
#   pull(Gene))
dplot_markers <- input[['dotplot_Gene']]
input[['dotplot_groups']] <- c('CellType_predict')
input[['dotplot_filter_cat']] <- 'CellType_predict'
input[['dotplot_filter_on']] <- (meta_filter %>%
                                   filter(grepl('Amacrine|Rod|Cone|Retinal|Muller|Horizon|Bipol|Astro|RPE', CellType_predict)) %>% 
                                   pull(CellType_predict) %>% unique())
d <- make_dotplot(input, scEiaD_2020_v01, meta_filter, cat_to_color_df)
dplot <- d  + facet_grid(vars(cluster), scale = 'free_y', space = 'free') + theme(
  strip.background = element_blank(),
  strip.text.y = element_blank()
)

input <- list()
input[['meta_column']] <- 'CellType_predict'
input[['pt_size_back']] <- 1
input[['pt_size_meta']] <- 1
#input[['gene_and_meta_scatter_tech']] <- 'Droplet'
input[['meta_column_transform']] <- 'None'
input[['label_toggle']] <- 2
input[['meta_filter_cat']] <- 'CellType_predict'
input[['meta_filter_on']] <- NULL
ctp <- make_meta_scatter_umap_plot(input, mf, meta_filter,
                                   celltype_predict_labels,
                                   celltype_labels,
                                   tabulamuris_predict_labels,
                                   cluster_labels,
                                   cat_to_color_df,
                                   velocity = TRUE
)

input <- list()
input[['meta_column']] <- 'CellType_predict'
input[['pt_size_back']] <- 4
input[['pt_size_meta']] <- 2
input[['gene_and_meta_scatter_tech']] <- 'Droplet'
input[['meta_column_transform']] <- 'None'
input[['meta_filter_cat']] <- 'CellType_predict'
input[['meta_filter_on']] <- c('AC/HC Precursors','Amacrine Cells','Bipolar Cells','Cones','Early RPCs','Endothelial','Horizontal Cells','Late RPCs','Microglia','Muller Glia','Neurogenic Cells','Pericytes','PR Precursors','Retinal Ganglion Cells','Rod Bipolar Cells','Rods','RPCs','Tabula Muris')
ctp2 <- make_meta_scatter_umap_plot(input, mf, meta_filter,
                                    celltype_predict_labels,
                                    celltype_labels,
                                    tabulamuris_predict_labels,
                                    cluster_labels,
                                    cat_to_color_df
)

input <- list()
input[['meta_column']] <- 'CellType_predict'
input[['pt_size_back']] <- 1
input[['pt_size_meta']] <- 1
#input[['gene_and_meta_scatter_tech']] <- 'Droplet'
input[['meta_column_transform']] <- 'None'
input[['meta_filter_cat']] <- 'CellType_predict'
input[['meta_filter_on']] <- NULL
org <- make_meta_scatter_umap_plot(input, mf, meta_filter,
                                   celltype_predict_labels,
                                   celltype_labels,
                                   tabulamuris_predict_labels,
                                   cluster_labels,
                                   cat_to_color_df
)

plot_grid(dplot + theme(legend.title = element_text(angle = 90),
                        axis.text = element_text(size = 6)),
          plot_grid(ctp$plot + theme(legend.position = "none"), 
                    ctp2$plot + theme(legend.position = "none") + facet_wrap(~CellType_predict, ncol = 4),
                    grid.grabExpr(draw(ct_confusion)),
                    org$plot + facet_wrap(~organism) + theme(legend.position = "none"), 
                    ncol = 2,
                    labels = c('b', 'c', 'd', 'e'),
                    rel_heights = c(1,0.8)), 
          ncol = 2, 
          rel_widths = c(1,3), 
          labels = c('a',NULL))

```

## ML cell type labels result in high study diversity for each cell type

After ML projection of cell type labels from the original `r meta_filter %>% filter(!is.na(CellType)) %>% nrow()` labels onto a total of `r meta_filter %>% filter(!is.na(CellType_predict) | !is.na(TabulaMurisCellType_predict)) %>% nrow()` cells we have substantially improved the number of studies per cell type. For example, we went from `r joinedSA %>% filter(CellType == 'Muller Glia') %>% pull(2)` human studies with labeled Müller Glia to `r joinedSA %>% filter(CellType == 'Muller Glia') %>% pull(5)` after labeling (`r supTab_cap(name = 'supTab_SAcounts', display = 'cite')`). Overall we go from an average of `r joinedSA %>% filter(CellType != 'Unlabelled') %>% pull(2) %>% mean() %>% round()` studies per human cell type to `r joinedSA %>% filter(CellType != 'Unlabelled') %>% pull(5) %>% mean() %>% round()` and `r joinedSA %>% filter(CellType != 'Unlabelled') %>% pull(4) %>% mean() %>% round()` studies per mouse cell type to `r joinedSA %>% filter(CellType != 'Unlabelled') %>% pull(7) %>% mean() %>% round()` after transferring the cell type labels.

As another check on the quality of the cell type assignments, we ran the cell and cluster independent haystack gene search and pairwise differential expression tests between the predicted cell types (see methods for further details). We show the five most differentially expressed genes for each of the major retina cell types are consistent with known retinal cell markers (`r fig_cap(name = 'figXGboost_umap', display = 'cite')`a). As a simple metric to identify known and unknown genes relating to the cell type specific expression we search PubMed for the number of publications with two searches per gene. We expect most of the genes identified to be known in the literature. The first search is the more precise "gene AND cell type" (e.g. "PDE6H AND Cones") and the second search is the more inclusive "gene AND retina" (e.g. "PDE6H AND Retina"). Of the 50 genes in (`r fig_cap(name = 'figXGboost_umap', display = 'cite')`a), `r 50 - top_markers %>% filter(Gene %in% dplot_markers) %>% filter(c1 == 0) %>% nrow` had one or more citations in the gene - cell type search (Supplementary File celltype_markers.tsv) and `r 50 - top_markers %>% filter(Gene %in% dplot_markers) %>% filter(c1 == 0, c2 == 0) %>% nrow` had one or more citation in the inclusive search. The 50 genes had a mean of `r top_markers %>% filter(Gene %in% dplot_markers) %>% pull(c2) %>% mean() %>% round()` studies (with the inclusive gene by "retina" search). In contrast, 100 randomly chosen genes had a mean of `r pmid3 %>% map(function(x) sum(!is.na(x))) %>% unlist() %>% enframe() %>% pull(value) %>% mean() %>% round` (wilcox test p < 1.44 x 10^-17).

<!-- (broom::tidy(wilcox.test(pmid3 %>% map(function(x) sum(!is.na(x))) %>% unlist() %>% enframe() %>% pull(value), top_markers %>% filter(Gene %in% dplot_markers) %>% pull(c2))))$p.value -->

## The scVI-based scEiaD UMAP projection blends batches and species while separating cell types

The 2D UMAP projection of the scVI-calculated batch corrected 8 latent dimensional space blends the `r meta_filter %>% pull(study_accession) %>% unique() %>% length() - 1` studies together while also maintaining distinct space for the `r meta_filter$CellType_predict %>% table() %>% length()` unique cell types. We also see good mixing across all the droplet and well based single cell technologies (`r supFig_cap(name = 'supFig_platformInt', display = 'cite')`). We see the neurogenic and progenitor populations from which the retinal cell types are derived near the center of the UMAP visualization. The photoreceptor precursors are adjacent to the neurogenic population and, as demonstrated by the RNA velocity dynamics, flow into the rods and cones. The amacrine and horizontal precursors (AC/HC) likewise flow from the neurogenic center into the mature amacrine and horizontal cells.

The photoreceptors (cones and rods) of the retina which are responsible for color and low-light vision, respectively, are near each other in the UMAP space. The major remaining retina cell types, by proportion in the mammalian eye are the Müller Glia, which are a glial cell type which help support the neurons of the retina. Next we have the neural cell types which transmit and help interpret the signals from the photoreceptors before they leave the retina via the optic stalk: the amacrine cells, retinal ganglia, horizontal, and bipolar cells. All of these cells are in well separated spaces in the UMAP. Finally we see across species that the major cell types overlap each other (`r fig_cap(name = 'figXGboost_umap', display = 'cite')`d). The macaque retina cells are not present in the precursor/neurogenic center of the UMAP as expected because only fully developed tissues were sampled in these data.

<!-- ## Pseudobulk testing leverages high biological diversity to provide robust differential expression results -->

<!-- Differential gene expression testing between clusters or cell types in the single cell field usually involve using running the statistical test of choice with all individual cells across the group. In bulk RNA-seq the tests are done in relation to replicates, which are generally required. As we have a large number of replicates, we can sum our gene expression counts by group (e.g. celltype or cluster, split by organism and study). After this summing, we now have created a "pseudo-bulk" matrix. The statistical properties (CITE SOMETHING) are now similar to the traditional bulk RNA-seq experiment, allowing us to leverage the robust tooling. We use the edgeR test to run our differential expression tests.  -->

While by eye the UMAP 2D projection generally blends together the three different species (macaque, mouse, human) in the UMAP visualization, we more rigorously tested this by tweaking the inputs to our xgboost cell type predictor machine learning system. Briefly, we trained two new models: one with only human data and one with only mouse data. We then applied this model to the other two species to see whether the model trained on one species was generally transferable to another species. Again, both models (human only and mouse only) both proved to be highly accurate at predicting cell types in the other species with the terminal cell types (`r supFig_cap(name = 'supFigML_CT_transfer', display = 'cite')`. 
## Projection of outside data onto scEiaD demonstrates similarities and differences of iPSC and organoids to primary cells

The hard work of creating this resource can be leveraged and extended by the wider community with a few relatively simple steps and modest compute requirements. Very briefly, if outside groups quantify their mouse or human scRNA with kallisto and the same references (see methods), they can overlay their data on top of scEiaD by 1. installing scVI (version 0.9.0 or higher), 2. downloading our 13 megabyte scVI model, and 3. following the Jupyter notebook on Google colab that we provide as a live demo available from https://github.com/davemcg/scEiaD. We demonstrate the power of this approach in two ways.

First, a wild-type retinal organoid dataset from Kallman et al was projected onto scEiaD [@kallmanInvestigatingConePhotoreceptor2020]. In the Google colab notebook we show how a subset of the reads from Kallman et al.'s SRR12130660 can be processed from the raw reads to a UMAP visualization in under 10 minutes. Indeed we see how organoid-based retinal cell types can be detected overlapping primary cells for RPCs, photoreceptors, amacrine cells, and retinal ganglion cells (`r supFig_cap(name = 'projection_fig', display = 'cite')`).

Finally, we demonstrate how the RPE derived from the Bharti iPSC differentiation process express canonical RPE markers of TTR and RPE65 but end up in a slightly different position in the UMAP projection`r fig_cap(name = 'RPE_fig', display = 'cite')`a). Differential expression analysis (see methods) identified that vimentin is nearly exclusively expressed in the iPSC-based RPE `r fig_cap(name = 'RPE_fig', display = 'cite')`b). This result is not surprising as Hunt et al. demonstrated that cultured and proliferating RPE express high levels of vimentin [@huntAlteredExpressionKeratin1990] and this same trend is seen in bulk RNA-seq from stem cell RPE and primary RPE `r supFig_cap(name = 'eyeIntegration_screenshot', display = 'cite')`) [@swamyEyeDiskEyeIntegration2019].

```{r RPE_fig, fig.width=14, fig.height=5,  fig.cap=RPE_fig_cap, echo=FALSE, message = FALSE, warning = FALSE}
RPE_fig_cap <- fig_cap(name='RPE_fig', caption =  'a. RPE distribution colored by study demonstrates how the most RPE are in two locations. iPSC-based RPE we provided are located more enriched in cluster 47. Tissue RPE more enriched in cluster 34 b. Violin plot of two functional RPE markers (TTR, RPE65) and vimentin (proliferating RPE marker)')
input <- list()
input[['pt_size_back']] <- 5
input[['pt_size_meta']] <- 10
input[['meta_column']] <- 'study_accession'
#input[['gene_and_meta_scatter_tech']] <- 'Droplet'
input[['meta_column_transform']] <- 'None'
input[['meta_filter_cat']] <- 'CellType_predict'
input[['meta_filter_on']] <- c('RPE')
rpe <- make_meta_scatter_umap_plot(input, mf, meta_filter,
                                   celltype_predict_labels,
                                   celltype_labels,
                                   tabulamuris_predict_labels,
                                   cluster_labels,
                                   cat_to_color_df
)

genes4 <- c('VIM (ENSG00000026025)', 'RPE65 (ENSG00000116745)', 'TTR (ENSG00000118271)')
gene_list_dat <- list()
for (i in genes4){
  gene_list_dat[[i]] <- 
    scEiaD_2020_v01 %>% 
    tbl('metadata_filter') %>% 
    filter(cluster %in% c('34','47'), CellType_predict == 'RPE') %>% 
    left_join(., scEiaD_2020_v01 %>% tbl('counts') %>% 
                filter(Gene %in% c(i))) %>% 
    collect() %>%
    mutate(Gene = case_when(is.na(counts) ~ i, TRUE ~ Gene), 
           counts = case_when(is.na(counts) ~ 0, TRUE ~ counts))
}
library(gghalves) 
cluster_coords <- meta_filter %>% filter(cluster %in% c('34','47'), CellType_predict == 'RPE') %>% group_by(cluster) %>% 
  mutate(cluster = case_when(cluster == '47' ~ '47 (iPSC RPE)', TRUE ~ '34 (Tissue RPE)')) %>% summarise(UMAP_1 = mean(UMAP_1), UMAP_2 = mean(UMAP_2))

rpe_plot <- rpe$plot + coord_cartesian(xlim = c(-11,-2), ylim = c(-15,-5)) +
  geom_label_repel(data = cluster_coords, aes(x=UMAP_1, y= UMAP_2, label = cluster), force = 300)
gene_list_dat %>% bind_rows() %>% group_by(cluster) %>% summarise(UMAP_1 = mean(UMAP_1), UMAP_2 = mean(UMAP_2))

vln_plt <- gene_list_dat %>% bind_rows() %>%  mutate(cluster = case_when(cluster == '34' ~ '34 (Tissue RPE)', TRUE ~ '47 (iPSC RPE)')) %>% ggplot(aes(x=Gene, y=counts, color = cluster))   + geom_half_boxplot() + geom_half_violin(side = 'r', aes(fill=cluster), scale = 'width') + cowplot::theme_cowplot() #+ facet_wrap(~organism)

plot_grid(rpe_plot ,
          vln_plt + coord_flip() + ggsci::scale_fill_nejm() + ggsci::scale_color_nejm(),
          labels = c('a','b'))
```

# Methods

## Reproduciblity

The set of Snakemake pipelines that takes in the raw fastq sequence and outputs the scEiaD is at https://github.com/davemcg/scEiad [@kosterSnakemakeScalableBioinformatics2012]. The publication commit is #9d3c5e1 Furthermore, the repository has been deposited at Zenodo under 10.5281/zenodo.4638117 We will briefly discuss the pipeline choices, programs and algorithms, and versions below. For the R packages, we provide package versions as the supplementary file "R_session_info.txt"

## Quantification of gene counts

Gene quantification is handled by the SnakeQUANT snakefile. First, we generated multiple quantification indices to facilitate calculation of RNA velocity. For each droplet technology, we generated a separate set of transcript sequences that contain both exonic and intronic sequences using the “get_velocity_files” function from the R package BUSpaRse. The reference transcriptome annotation used to build sets of transcript sequences were the Gencode “gencode.vM25.annotation.gtf.gz” and “gencode.v35.annotation.gtf.gz” for mouse and human, respectively [@harrowGENCODEReferenceHuman2012; @jGENCODEReferenceHuman2012]. Because the Macaca Fascicularis genome is less well annotated we used the Ensembl release 101  genome and transcriptome annotation “Macaca_mulatta.Mmul_101.gtf.gz”  [@yatesEnsembl20202020].  A single set of exonic transcript sequences was created for each species for all well techonologies. A kallisto quantification index was generated for each of these fastas using kallisto index (0.46.2) [@brayNearoptimalProbabilisticRNAseq2016]. Well based samples were quantified using kallisto quant. For the relatively few single ended samples, the params “–single -l 200 -s 30” were used for kallisto quant. Otherwise the “–bias” flag was added. For the droplet-based samples, we adapted the bustools workflow for generating spliced and unspliced count matricies. (0.39.4) [@melstedBarcodeUMISet2019].

## Intersection of gene names between mouse, macaque, and human

To facilitate comparison of gene expression across species, where possible we converted mouse and macaque gene ids and names to human ones. We downloaded a mapping of orthologous genes between human, mouse, and macaque using the Ensembl BioMart web browser in November 2020. We identified 15,759 human genes that could be directly mapped to mouse and macaque orthologs. Genes present in mouse or macaque that were not found in human were not used for HVG gene selection, but were retained and used for differential gene expression.

## Custom macaque reference quantification

As we noticed that several retina marker genes (e.g. NRL and CRX) had very low expression in the macaque data we quantified the scRNA data twice: once with the Ensembl reference and again with the same Gencode human reference used for the human data. We compared the gene-level counts for each cell and replaced the macaque gene count with the human counts if the human counts were greater than the macaque counts and, to prevent genes with very few total counts from being used, we required the counts greater than the first quartile of non-zero macaque gene expression.

## Remove empty droplets and further QC.

After bustools count, we used R (3.6.2) to remove empty droplets. The BUSpaRse package was used to input the bustools counts mtx file. The DropletUtils package with the "barcodeRanks" function was used to automatically detect the inflection point in the barcode count ranks that delineates the likely empty droplets [@lunEmptyDropsDistinguishingCells2019]. We then removed cells with percent mitochondrial reads of >10%. After merging the individual count matrices into one sparse matrix, we created a Seurat version 3 object and removed cells with fewer than 200 detected unique genes, and for the droplet data, more than 3000 detected genes (these are likely to be doublets) [@butlerIntegratingSinglecellTranscriptomic2018].

## Normalization and batch effect correction

The following steps (normalization through benchmarking are handled by the SnakePOP file) We tested several gene count normalization approaches as we were not certain which would produce an optimal outcome: standard (default Seurat, library size normalization, then log transform), sqrt (same, but with sqrt normalization), libSize (omit the log or sqrt normalization), scran, SCT from Seurat, and for scVI, no normalization (counts) [@lunStepbystepWorkflowLowlevel2016]. Our R implementation of the normalization approaches as well as how we constructed the Seurat v3 object can be found in the supplementary file "make_seurat_obj_functions.R"

## Batch normalization under a grid search procedure

We tested `r methods` against 2000 HVGs, the different gene count normalization procedures discussed above, and both 8 and 30 outputted batch corrected latent dimensions. The latent dimensions are the input for clustering, the 2D UMAP visualization, and the xgboost machine learning to transfer cell type labels to unlabelled cells. We were unable to run every method successfully with every normalization method. Magic could not complete with the standard or libSize normalization. CCA could not complete with the libSize normalization. We also tried the DESC, liger, and Conos batch corrections methods but were unable to get them to work reliably so they were dropped. The batch correction step implementation can be found the supplementary file "merge_methods.R" and in the github repo (https://github.com/davemcg/scEiaD/blob/master/src/merge_methods.R).

## Clustering and UMAP
Louvain-Jaccard clustering against the batch corrected latent dimensions used the Seurat implementation [@blondelFastUnfoldingCommunities2008]. We tried the k-nearest neighbors (knn) parameters 5 and 7 (where 5 gives more clusters than 7). We also used the leiden algorithm as implemented by PARC with a resolution of 0.6 and 0.8 (higher results in more clusters) [@traagLouvainLeidenGuaranteeing2019; @stassenPARCUltrafastAccurate2020]. These two resolutions were chosen as they roughly gave the same number of clusters at the Seurat Louvain-Jaccard approach with a knn 7.

The UMAP visualization was calculated with the Seurat "RunUMAP" using the uwot R package [@mcinnesUMAPUniformManifold2020]. We tried min.dist parameters of 0.001, 0.1, and 0.3 and tried n.neighbors across 15, 30, 100, and 500. A smaller min.dist value gives "tighter" groupings while a higher number of n.neighbors uses a larger number of near cells to calculate the global positioning.

## Benchmarking and scPOP

We wrote the scPOP R package to unify the LISI and Silhouette metrics from Harmony and kBet, respectively, along with NMI and ARI [@buttnerTestMetricAssessing2019; @korsunskyFastSensitiveAccurate2019]. LISI and Silhouette require a dense matrix, which is a problem for our data as a `r meta_filter %>% nrow()` cell by 8 latent dimension dense matrix cannot fit in out largest available memory node (1.5 TB). We down-sampled the dataset to ~100,000 cells, taking care to keep all rarer cell types for the LISI and Silhouette benchmarking.

To merge these metrics into a balanced single score, we Z scale each and sum them. scPOP produces both tables and visualizations allowing the user to quickly see both the interplay of batch mixing and cluster/cell type separation and the overall performance. If a user wishes to prioritize batch mixing or cluster/cell type separation we let the user provide a custom batch/cluster-cell type scaling value (1 is the default).

## Multi-step doublet removal
To identify probable doublets (more than one cell in a droplet) we ran DoubletDetect and scrublet and calculated the distribution of DoubletDetect and scrublet scores across all clusters and removed clusters with a score in both metrics greater than 4 standard deviations above the mean [@adamgayosoJonathanShorDoubletDetectionDoubletdetection2020; @wolockScrubletComputationalIdentification2019]. This removed another `r format((umapRef %>% nrow() %>% as.integer()) - (meta_filter %>% nrow() %>% as.integer()), big.mark=",", scientific=FALSE)` cells, leaving `r format((meta_filter %>% nrow()) %>% as.integer(), big.mark=",", scientific=FALSE)` in total.

## xgboost based cell type model

In order to identify cell types for the `r format(meta_filter %>% filter(is.na(CellType), is.na(TabulaMurisCellType)) %>% nrow(), big.mark=",", scientific=FALSE)` unlabelledcells we designed a custom xgboost based cell type classifier. We took labeled data and split it into training (2/3) and test (1/3) sets, stratified by cell type. The input features used to train the model are the scVI latent dimensions, the total number of reads in each cell, the number genes detected in each cell, and the percent mitochondrial gene expression of each cell. We additionally generated features using the age of each sample by group sample into three developmental categories (Early Development, Late Development, and Adult) and then generated a one-hot encoded feature for each category. In order to speed up training times, we used the gpu implementation of the xgboost algorithm from the the xgboost python library. The model was trained using default parameters. The trained model had an overall macro and micro AUC score of 0.98 and 0.99, respectively. This model was then used to identify labels for each all cell. Unlabeled data was pre-processed identically to training data and fed into model to generate a vector of label probabilities for each cell. We selected the highest label probability for each cell, and required a minimum probability of 0.5 to assign a label to a cell.

For the organism specific xgboost ML we followed the above procedure, except that we combined Early RPCs, Late RPCs, and RPCs into one category and did not attempt to predict the non-retina cell types (e.g. fibroblasts) as there were very few labelled cells across all three organisms.

<!-- ## pseudobulk differential expression testing -->

<!-- Our pseudo-bulk differential testing was inspired by the OSCA guide. We computed differential tests for: cell type (published/transferred) against remaining cell types, celltype (published/transferred) pairwise against other individual celltype (published/transferred), specific cluster against remaining clusters, specific cluster against different individual cluster (pair-wise). Briefly, we grouped cells by the categories (e.g. each unique celltype (transferred) - study combination) and summed all gene counts. After summing the scores, we built a model matrix as 0 ~ categories + organism to remove any species specific effects. When then used the edgeR "estimateDisp" and tested the fit with "glmQLFit" and the parameter "robust = TRUE". Specific contrasts (e.g. Rods vs Cones) were differentially tested with "glmQLFTest". -->

## Marker gene identification

To identify marker genes across the CellType (predict) and cluster groups, we used the scran findmarkers (wilcox test) along with the singleCellHaystack algorithm. The scran findmarkers test runs a wilcox test in a pairwise manner (e.g. Rods vs all other cell types). It returns an overall p-value (and FDR) that assesses how well the gene is at separating the group of interest from all other cells. It also returns for each pair-wise comparison an area under the curve (AUC) score, where 1 is a perfect power to distinguish and 0 is no power. The singleCellHaystack algorithm uses a Kullback-Leibler divergence (D KL) measurement of the scVI lower dimensional space to identify genes with non-random distribution. A higher D KL score represents a gene with "specific" expression in the lower dimensional space and is used to calculate a FDR corrected p value against the full distribution of D KL values. We filtered to keep genes with scran FDR < 1, a mean AUC > 0.2, and a log10(D KL FDR) < -10000.  No more than 50 genes for each cell type were retained (sorted by mean AUC).

<!-- ## Trajectory -->

<!-- To calculate continuous trajectories for individual terminally differentiated retina cell types, we used the Slingshot R package [@streetSlingshotCellLineage2018] on the human subset. Slingshot first uses the clusters to build a minimum spanning tree to identify the global lineage structure. We specified the starting cluster by identifying the cluster for human and mouse with a high proportion of the early RPC cells and specified the end clusters that were predominantly a terminal retina cell type (Amacrine, (Rod) Bipolar, Cones, Horizontal, Muller Glia, Retinal Ganglion, Rods). As some clusters were combinations of different (but similar) cell types, we "subclustered" the clusters by splitting them into smaller groups if they contained two or more cell types with > 20% ratio of a cell type. XXXX trajectories were built and we hand-pruned out redundant or non-sensical (e.g. Photoreceptor precursors to Cone to Rod Bipolar) trajectories, leaving XXXX trajectories for downstream analysis.  -->

<!-- To identify major points of gene expression changes during differentiation we built linear models across XXXXXXX bins.  -->

## Calculation of RNA velocity

RNA velocity calculations were with the velocriaptor wrapping of the scVelo python library [@bergenGeneralizingRNAVelocity2020]. From the anndata objects generated by our Snakemake pipeline we calculated velocity across all genes. Genes without detectable velocity were dropped. The scVI generated latent space (instead of PCA) was used to calculate first and second order moments. The calculated moments were used to estimate RNA velocity. Differential velocity was tested between celltypes using pairwise wilcox rank sum tests.

<!-- To id -->

# Conclusion

# Limitations

The scVI model is first built the on human data. The mouse and macaque data are then projected (or queried) onto it with the scVI implementation of the scArches method. While this system works very well to integrate information between these three species, this approach may not scale to more distantly related species. Another limitation is that discrepancies between cell type labels between different labs makes certain transitioning cell type labels a bit imprecise. One example is how the rods and photoreceptor precursor labels partially overlap. Though we attempted to ameliorate the issue by removing cell type labels in large disagreement with the consensus, some disagreements could propagate into our machine labeled cells type assignments. These issues may reflect labeling continuous processes with discrete labels.

While scEiaD distinguishes the major cell types very well, some of the cell types contain many "sub types" - notably the Amacrine cells have a huge variety in morphology with a recent study identifying over 60 different types of Amacrine cells in mouse. At this time our batch corrected pan retina cell space does not precisely resolve these sub cell types with high resolution. We are actively working to "sub cluster" the cell types so we can robustly and reliably identify the high diversity of retinal cell types across the entire retina.

## The scEiaD is a unique ocular resource that provides a highly diverse, large N dataset with a relatively small amount of compute power

We have assembled the largest ocular single cell transcriptome database to date. The rapid of advancement of algorithms to batch correct and process data continue to reduce the computational requirements to handle huge numbers of cells. The scVI batch correction step on around one million cells runs within a few hours on a GPU and 150GB of memory. This places this crucial step within the capabilities of a moderately powerful computer or a cloud compute node. Further downstream processing can largely be done a computers with 64+ Gb of memory and a few hundred GB of disk space. We believe that our efforts can be replicated in any other tissue / system with a large number of independent studies by a small number of computational scientists following our general approach. We provide the completed analysis as both Scanpy (h5ad) and Seurat objects at https://github.com/davemcg/scEiaD.

## Benchmarking and quantitation of integration performance is crucial for meta-atlas studies

We originally intended to use the Seurat CCA method to integrate the datasets. However, the long run-times of CCA and poor integration of our known cell types led us to benchmark more methods and parameters. After adding more integration methods we first attempted "hand-assess" the integration results by using the UMAP 2D projection view. This proved to scale poorly and this led use to curate some of the more useful benchmarking algorithms (NMI, ARI, silhouette, and LISI) that roughly matched our "hand-assessed" results into the scPOP R package. While we chose the scVI algorithm, we strongly suggest any other groups attempting a similar meta-atlas construction chose a quantifiable set of criteria so optimal methods and parameters can be picked.

## Transfer of cell type labels from a smaller number of studies onto the remainining cells is a powerful way to increase diversity in a meta-atlas

We first hand curated over 350,000 published cell type labels across 24 publications. With a xgboost algorithm using the latent dimensions, cell age classification (developing or matured), and the UMAP coordinates, we can very accurately label the remaining cells. Many other cell type labeling algorithms and systems exist for those groups less willing or able to tune a machine learning algorithm. For example, the developers of scVI also have a cell type label projection algorithm called scANVI. Whichever approach you use, taking a smaller number of high quality labels and projecting them onto the remaining cells is a powerful way to leverage community knowledge across a huge diverse dataset.

## Projection allows community knowledge to be leveraged by all

Many retina atlases have been published to date. We argue that we have created the first atlas that is generally useful because 1. our dataset/atlas is several times larger than any other published set, 2. our data is available via download in several forms at https://github.com/davemcg/scEiaD, and crucially 3. we provide a Google colab/Jupyter notebook which exactly lays out how to use scVI to project (or query) outside data onto our scEiaD with minimum compute resources. We demonstrate concretely how this can work by showing how iPSC RPE can be queried onto the reference dataset to demonstrate both similarites and dissimilarities in their transcriptomes.

# Supplemental Information
## Tables
```{r supTab1_CTcounts, fig.width=4, fig.height=14, out.width=300, echo = F, results='asis', warning = F, fig.cap = supTab1_CTcounts_cap, fig.pos = 'H'}
supTab1_CTcounts_cap <- supTab_cap(name='supTab1_CTcounts', caption =  'Counts for cell type labels. Published are the author created labels from the published datasets. Transferred are the cell labels that were transferred by our xgboost-based machine learning model onto the entire scEiaD dataset.')
cat("<div id=\"tbl:supTab1_CTcounts\">")
source('figs_and_tables/sup_fig__xgboost_PR.R')
source('figs_and_tables/sup_fig2__celltype_xgboost_labelling.R')
fontsize(ctTable  %>% padding(padding = 0, part = "body") %>% align(align = 'center', part = 'all') %>% autofit(), size = 8, part = 'all')
#Output Caption (* * for italics)
cat(paste0("\n*",supTab1_CTcounts_cap,"*\n"))
# Output </div>
cat("</div>")
```

```{r supTab_AUCxgboost, fig.cap=supTab_AUCxgboost_cap, fig.width=4, echo=FALSE, results='asis', message = FALSE, warning = FALSE}
supTab_AUCxgboost_cap <- supTab_cap(name='supTab_AUCxgboost', caption =  'Area under the precision recall curve (AUC) for each cell type, split by study. The "All" study is the AUC score across all cells within the cell type')

cat("<div id=\"tbl:supTab_AUCxgboost\">")
fontsize(xgboost_pr_table_full %>% arrange(`Cell Type`, Study) %>% flextable() %>%  padding(padding = 0, part = "body") %>% align(align = 'center', part = 'all') %>% autofit(), size = 8, part = 'all') %>% colformat_double(digits = 2)
#Output Caption (* * for italics)
cat(paste0("\n*",supTab_AUCxgboost_cap,"*\n"))
# Output </div>
cat("</div>")
```

```{r supTab_CTmislabels, fig.width=2, fig.cap=supTab_CTmislabels_cap, echo=FALSE, results='asis', message = FALSE, message = FALSE, warning = FALSE}

supTab_CTmislabels_cap <- supTab_cap(name='supTab_CTmislabels', caption =  'Counts of cell type labels with our xgboost machine learning system (PredCellType) and the published cell type labels (TrueCellType). Ratio is calculated as CellType that were labelled as CellType_predict. Ratio < 0.05 were filtered out from view in the table. ')

cat("<div id=\"tbl:supTab_CTmislabels\">")
#source('figs_and_tables/sup_fig2__celltype_xgboost_labelling.R')
fontsize(meta_filter %>% filter(!is.na(CellType_predict), !is.na(CellType), is.na(TabulaMurisCellType_predict)) %>% 
           mutate(pred_correct  = ifelse(CellType == CellType_predict, 'Correct', 'incorrect')) %>% 
           group_by(CellType, CellType_predict) %>% 
           summarise(Count = n()) %>% 
           mutate(Ratio  = round(Count / sum(Count),3 )) %>% 
           filter(Ratio > 0.02) %>% 
           filter(!grepl('Doubl|Periocu', CellType)) %>% 
           ungroup() %>% 
           group_by(CellType, CellType_predict) %>% 
           arrange(CellType, -Ratio) %>% 
           flextable()  %>% 
           padding(padding = 0, part = "body") %>% align(align = 'center', part = 'all') %>% autofit(), 
         size = 8, 
         part = 'all') %>% 
  colformat_double(digits = 2)
#Output Caption (* * for italics)
cat(paste0("\n*",supTab_CTmislabels_cap,"*\n"))
# Output </div>
cat("</div>")
```

```{r supTab_SAcounts, fig.width=8, fig.height=14, out.width=300, fig.cap=supTab_SAcounts_cap, echo=FALSE, results='asis', message = FALSE, warning = FALSE}
cat("<div id=\"tbl:supTab_SAcounts\">")
supTab_SAcounts_cap <- supTab_cap(name='supTab_SAcounts', caption =  'Counts for number of studies with cell types labels before and after cell type label transfer')

fontsize(saTable  %>% padding(padding = 0, part = "body") %>% align(align = 'center', part = 'all') %>% autofit(), size = 8, part = 'all') 
#Output Caption (* * for italics)
cat(paste0("\n*",supTab_SAcounts_cap,"*\n"))
# Output </div>
cat("</div>")
```

## Figures

```{r supFig1, fig.width=4, fig.height=6, out.width=300, fig.cap=supFig1_cap, echo=FALSE, message = FALSE, warning = FALSE}
supFig1_cap <- supFig_cap(name='supFig1', caption =  'The left bar delineates the number of cells for each organism - technology combination. The right bar specifies the number of each cells in each post QC category. In silico doublets were identified with scrublet and DoubletDetector.')

source('figs_and_tables/sup_fig1__qc_failure.R')
supFig1_plot
```

```{r supFig_benchmark, fig.width=12, fig.height=8, out.width=300, fig.cap=supFig_benchmark_cap, echo=FALSE, message = FALSE, warning = FALSE}
supFig_benchmark_cap <- supFig_cap(name='supFig_benchmark', caption =  'Performance of the various batch correction tools across various benchmarking metrics. For the LISI and Silhouette plots in A, B higher (y-axis) means better batch mixing and further to the right (x-axis) means better cluster purity. For the ARI and NMI metrics (which reflects how well cluster matches with cell type) in C, D, higher means a better score.')

source('figs_and_tables/fig2__integrationPerf.R')
cowplot::plot_grid(
  cowplot::plot_grid(lisi + theme(legend.position="none"), silhouette, NULL, rel_widths = c(1,1,0.2), nrow = 1, labels = c('a' ,'b')), 
  cowplot::plot_grid(ari, nmi, legend, rel_widths = c(1,1, 1, 0.2), ncol = 4, labels = c('c', 'd')), 
  nrow = 2)
```

```{r supFigShekharSMART, fig.width=3, fig.height=3,  fig.cap=supFigShekharSMART_cap, echo=FALSE, message = FALSE, warning = FALSE}
supFigShekharSMART_cap <- supFig_cap(name='supFigShekharSMART', caption =  'Our xgboost ML properly labels this Shekhar et al. RBC FAC sorted population as enriched in RBC')
screenshot_paper <- ggdraw() + draw_image("figs_and_tables/shekhar_screenshot.png", scale = 1.2)

shekhar_counts <- meta_filter %>% 
  filter(study_accession == 'SRP073242', !is.na(CellType_predict)) %>% 
  group_by(CellType_predict) %>% summarise(Count = n()) %>% 
  ggplot(aes(x=CellType_predict, y = Count, label = Count)) + geom_bar(stat = 'identity') + 
  geom_text(color = 'gray', size = 5) + 
  cowplot::theme_cowplot(margin(t = 5, r = 5, b = 5, l = 5, unit = "pt")) + 
  coord_flip()
cowplot::plot_grid(screenshot_paper, shekhar_counts, labels = c('a','b'), ncol =1)
```

```{r supFig_PR, fig.width=12, fig.height=10,  fig.cap=supFig_PR_cap, echo=FALSE, message = FALSE, warning = FALSE}
supFig_PR_cap <- supFig_cap(name='supFig_PR', caption =  'Precision recall curves for our xgboost cell type predictor model across each cell type predicted')

pr_split_curves
```

```{r supFig_platformInt, fig.width=12, fig.height=10,  fig.cap=supFig_platformInt_cap, echo=FALSE, message = FALSE, warning = FALSE}
supFig_platformInt_cap <- supFig_cap(name='supFig_platformInt', caption =  'Distribution of cell types across the 2D UMAP confirms that the cell types are being properly placed despite the wide variety of single sequencing platforms present.')
input <- list()
input[['meta_column']] <- 'CellType_predict'
input[['pt_size_back']] <- 5
input[['pt_size_meta']] <- 7
input[['gene_and_meta_scatter_tech']] <- 'Droplet'
input[['meta_column_transform']] <- 'None'
#input[['meta_filter_cat']] <- 'CellType_predict'
#input[['meta_filter_on']] <- 'Bipolar Cells'
ctp3 <- make_meta_scatter_umap_plot(input, mf, meta_filter,
                                    celltype_predict_labels,
                                    celltype_labels,
                                    tabulamuris_predict_labels,
                                    cluster_labels,
                                    cat_to_color_df
)

ctp3$plot + facet_wrap(~Platform)
```

```{r supFigML_CT_transfer, fig.width=10, fig.height=20,  fig.cap=supFigML_CT_transfer_cap, echo=FALSE, message = FALSE, warning = FALSE}
supFigML_CT_transfer_cap <- supFig_cap(name='supFigML_CT_transfer', caption =  'F1 scores (1 is perfect) of cell type prediction when using a human or mouse based xgboost cell type prediction model on other organisms (human, macaque, mouse).')

bind_rows(mouse_model__scoring %>% mutate(model = 'Mouse'), human_model__scoring %>% mutate(model = 'Human')) %>% filter(name != 'AC/HC_Precurs') %>% ggplot(aes(x=name,y=value, fill = model, label = round(value, 3))) + geom_bar(stat='identity', position = position_dodge2()) + facet_wrap(~organism, ncol = 1) + coord_flip() + geom_text(position = position_dodge2(width = 1)) + cowplot::theme_cowplot() + ylab('F1 Score') + xlab('Cell Type')
```

```{r projection_fig, fig.width=12, fig.height=10,  fig.cap=projection_fig_cap, echo=FALSE, message = FALSE, warning = FALSE}
projection_fig_cap <- supFig_cap(name='projection_fig', caption =  'Screen shot of Google colab notebook that demonstrates how to integrate external data into the scEiaD resource. We see in the screenshot how the organoid dataset contains many of the retinal cell types.')

ggdraw() + draw_image("figs_and_tables/colab_screenshot2.png", scale = 0.7)

```

```{r eyeIntegration_screenshot, fig.width=12, fig.height=10,  fig.cap=eyeIntegration_screenshot_cap, echo=FALSE, message = FALSE, warning = FALSE}
eyeIntegration_screenshot_cap <- supFig_cap(name='eyeIntegration_screenshot', caption =  'Screenshot of eyeIntegration bulk RNA-seq meta-analysis of vimentin expression in different RPE tissue sources')

ggdraw() + draw_image("figs_and_tables/rpe_eyeintegration.png", scale = 1)
```

# Works Cited